The Anatomy of a Temporal Summarization System

### 2015-02-27 experiments.tex
Latest rev ck

\[ \mathrm{\mathbb{E}[Gain]} = \frac{|S_n|}{|S|}% \;\;\; \mathrm{Recall} = \frac{|S_n|}{|N|}
\] 
where $S$ is the set of system updates, 
$S_n$ is the set of nuggets contained in $S$, and $|\cdot|$ is the number of
elements in the set.
To compute the set $S_n$ we match each system update to 0 or more nuggets, 
where an update matches a nugget if their semantic similarity is above 
a threshold. $S_n$ results from the unique set of nuggets matched.
Because an update can map to more than one nugget, it is possible to receive an
expected gain greater than 1; values greater than one can be thought of as a 
compression ration, i.e. an expected gain of 10 indicates the average update
contains 10 unique nuggets on average and that the summary is a 10~:~1 
compression of the nugget information.
An expected gain of 1 would indicate that every
sentence was both relevant and contained a unique piece of information.

@@

\fdcomment{consider removing AP+SalienceTR}
\textbf{AP+Salience Time Ranked (AP+SalienceTR)} Our final model is identical 
to the
AP+Salience model except at the novelty filter stage.
Within the novelty filtering stage, 
AP+Salience inspects exemplars in order of decreasing salience when computing
similarity to previous updates. Because the order in which we add updates
affects future similarity calculations, we evaluate an alternate ordering 
scheme (AP+SalienceTR) where exemplars are selected in order of increasing 
time since event start. The purpose of this model is to give preference to 
more recent updates, since timelyness is one of the desirable traits of our update
summaries.

>>>>

\[ \mathrm{\mathbb{E}[Gain]} = \frac{|S_n|}{|S|}% \;\;\; \mathrm{Recall} = \frac{|S_n|}{|N|}
\] 
where $S$ is the set of system updates, 
$S_n$ is the set of nuggets contained in $S$, and $|\cdot|$ is the number of
elements in the set.
To compute the set $S_n$ we match each system update to 0 or more nuggets, 
where an update matches a nugget if their semantic similarity is above 
a threshold. $S_n$ results from the unique set of nuggets matched.
Because an update can map to more than one nugget, it is possible to receive an
expected gain greater than 1.
%values greater than one can be thought of as a 
%compression ration, i.e. an expected gain of 10 indicates the average update
%contains 10 unique nuggets on average and that the summary is a 10~:~1 
%compression of the nugget information.
An expected gain of 1 would indicate that every
sentence was both relevant and contained a unique piece of information.

@@

%\fdcomment{consider removing AP+SalienceTR}
%\textbf{AP+Salience Time Ranked (AP+SalienceTR)} Our final model is identical 
%to the
%AP+Salience model except at the novelty filter stage.
%Within the novelty filtering stage, 
%AP+Salience inspects exemplars in order of decreasing salience when computing
%similarity to previous updates. Because the order in which we add updates
%affects future similarity calculations, we evaluate an alternate ordering 
%scheme (AP+SalienceTR) where exemplars are selected in order of increasing 
%time since event start. The purpose of this model is to give preference to 
%more recent updates, since timelyness is one of the desirable traits of our update
%summaries.
%

### 2015-02-27 feature_table.tex
Latest rev ck

\begin{figure}[t!]
\begin{tabular}{| l |} 
\hline
\textbf{Basic Features}\\
$\cdot$ Sent. position (normalized by doc. length) \\
$\cdot$ Sent. length \\
$\cdot$ Ratio of punc. to non-punc. chars. \\
$\cdot$ Ratio of caps. to non-caps. chars. \\
$\cdot$ Ratio of lowercase to other chars. \\
$\cdot$ Ratio of uppercase to other chars. \\
$\cdot$ \# of caps. words (normalized by \# of words)\\
$\cdot$ \# of Person, Location, Org. Date, Number,\\
$\;\;$ Ordinal, Percent, Money, Set, Misc  N.E. \\
$\;\;$ tags (normalized by \# of words)\\
\hline
\textbf{Query Features}\\
$\cdot$ \% of query words covered by sent.\\
$\cdot$ Total query matches.\\
$\cdot$ Total event-type synonyms/hypernyms/ \\
$\;\;$ hyponyms coverage.\\
$\cdot$ Total event-type synonyms/hypernyms/ \\
$\;\;$ hyponyms matches.\\
\hline
\textbf{Language Model Features}\\
$\cdot$ Avg. token log probability (domain lang. \\
$\;\;$ model)\\
$\cdot$ Avg. token log probability (background \\
$\;\;$ lang. model)\\
\hline
\textbf{Geo-tag Features}\\
$\cdot$ Median document distance to nearest  \\
$\;\;$ location cluster (current hour).\\
$\cdot$ Median document distance to nearest  \\
$\;\;$ location clusters (previous hour).\\
$\cdot$ Distance of first location in doc. to nearest \\
$\;\;$ location cluster (current hour).\\
$\cdot$ Distance of first location in doc. to nearest \\
$\;\;$ location cluster (previous hour).\\
\hline
\textbf{Temporal Features}\\
$\cdot$ Avg. tf-idf at current time.\\
$\cdot$ Change in avg. tf-idf since previous hour \\
$\;\;$  (up to 24 hours).\\
$\cdot$ Time since query/event start.\\
\hline
\end{tabular}
\caption{Salience model features.}
\end{figure}

>>>>

\begin{figure}[t!]
\begin{tabular}{| l |} 
\hline
\textbf{Basic Features}\\
$\cdot$ Sent. position (normalized by doc. length) \\
$\cdot$ Sent. length \\
$\cdot$ Ratio of punc./caps./lower/upper to \\
$\;\;$ other chars. \\
%$\cdot$ Ratio of caps. to non-caps. chars. \\
%$\cdot$ Ratio of lowercase to other chars. \\
%$\cdot$ Ratio of uppercase to other chars. \\
$\cdot$ \# of caps. words (normalized by \# of words)\\
$\cdot$ \# of Person, Location, Org., etc.\\
$\;\;$ (normalized by \# of words)\\
%$\;\;$ Ordinal, Percent, Money, Set, Misc  N.E. \\
\hline
\textbf{Query Features}\\
$\cdot$ \% of query words coverage/matches\\
$\cdot$ Total event-type synonyms/hypernyms/ \\
$\;\;$ hyponyms coverage/matches.\\
%$\cdot$ Total event-type synonyms/hypernyms/ \\
%$\;\;$ hyponyms matches.\\
\hline
\textbf{Language Model Features}\\
$\cdot$ Avg. token log probability (domain lang. \\
$\;\;$ model)\\
$\cdot$ Avg. token log probability (background \\
$\;\;$ lang. model)\\
\hline
\textbf{Geo-tag Features}\\
$\cdot$ Median document distance to nearest  \\
$\;\;$ location cluster (current hour/prev. hour).\\
%$\cdot$ Median document distance to nearest  \\
%$\;\;$ location clusters (previous hour).\\
$\cdot$ Distance of first location in doc. to nearest \\
$\;\;$ location cluster (current hour/prev. hour).\\
%$\cdot$ Distance of first location in doc. to nearest \\
%$\;\;$ location cluster (previous hour).\\
\hline
\textbf{Temporal Features}\\
$\cdot$ Avg. tf-idf at current time.\\
$\cdot$ Change in avg. tf-idf since previous hour \\
$\;\;$  (up to 24 hours).\\
$\cdot$ Time since query/event start.\\
\hline
\end{tabular}
\caption{Salience model features.}
\end{figure}

### 2015-02-27 methods.tex
Latest rev ck

% Updates should not only be relevant to the event but also salient,
% i.e. worthy of
% reporting.
% For most domains, updates also need to be
% timely and novel, i.e. users need to be informed of changes in the
% event as quickly as possible.
% Finally, in aggregate the updates must comprehensively summarize the
% event.

\begin{figure}
\setlength{\fboxsep}{10pt}
    \fbox{\parbox{6.9cm}{
        -- hurricane force wind warnings are in effect from Rhode Island Sound to Chincoteague Bay

        -- Obama declared an emergency in Maryland and signed an order authorized the Federal Emergency Management Agency to aid in disaster relief

        -- over 5000 commercial airline flights scheduled for October 28 and October 29 were cancelled
 }}
 \caption{Example nuggets from Hurricane Sandy.\label{fig:nuggets}}
\end{figure}

Throughout our treatment of our algorithm, the \emph{salience} 
of an update captures the degree to which it reflects an 
event's nuggets.  Assuming that we have a text representation 
of our nuggets, the salience of an update $u$  with respect to a set of nuggets $N$ is
defined as,
\begin{align}
\operatorname{salience}(u) = \operatorname{max}_{n \in N} 
\operatorname{sim}(u, n) \ref{eq:salience}
\end{align}
where $\operatorname{sim}(\cdot)$ is the semantic similarity such as
the cosine similarity of latent vectors associated with the update and 
nugget text \cite{?}. % We train
% a regression model to predict salience using features derived from
% the updates themselves
% and the document stream.

     % \ref{sec:methods}.  In our version of this problem, we assume that that
     % the system receives one batch of new sentence-segmented documents every
     % hour throughout the period of interest.
\subsection{Update Summarization}

Our system architecture follows a simple pipeline design where each
stage provides an additional level of processing or filtering of the input
sentences.
We begin with an empty update summary $U$.
At each hour we receive a new batch of sentences $b_t$ from the stream
and perform the following actions:
\begin{enumerate}
    \item predict the salience of sentences in $b_t$ (Section~\ref{sec:salpred}),
  \item select a set of exemplar sentences  in $b_t$ by combining clustering with 
      salience predictions (Section~\ref{sec:exsel}),
  \item add the most novel and salient exemplars to $U$ (Section~\ref{sec:upsel}).
\end{enumerate}

The resultant list of updates $U$ is our summary of the event.
% Step 2 uses the affinity propagation algorithm to incorporate the salience
% predictions from step 1. We refer to this summarization
% system as the \textbf{AP+Salience} model.

%\kmcomment{This was as best as I could do on this. These are the example
%    categories from the wikipedia page. However you refer in the language model
%    description to the event type earthquake which is more specific. Also, in
%    the TREC data you refer to event types that are specified in the metadata.
%So I'm wondering how many different categories you have and where they come
%from.}

%\fdcomment{include description of nuggets here.} 
%%\kmcomment{We should jointly
%    discuss what goes in data and what goes in problem definition. I have moved
%    event types here because I think knowing how many types and examples of
%    them would be helpful upfront. I think where nuggets and wikipedia pages go
%is questionable.}

%Figure \ref{alg:temporal-summarization} outlines our general update
%summarization algorithm.  At each hour, the system processes each input 
%sentence batch $S_t$.
%We predict the salience $P$ for all input sentences $s\in S_t$.
%Next, we cluster $S_t$ using the AP clustering algorithm, biased by $P$,
%obtaining a set of exemeplar sentences $E$. Finally we select a subset of $E$
%to be updates and add those to our set of summary updates $U$.
%
%\begin{algorithm}%[H]
% \KwData{ 
%     $S_{\stime}, S_{\stime+1},\ldots,S_{\etime} $ --- time ordered sentence
%     batches\\
% \KwResult{U --- a list of updates, i.e. the update summary} 
%}
% ~\\
% Initialize empty list $U$ of updates\\
%    \For{$t \gets \stime,\ldots,\etime $}{
%        $P \gets \operatorname{PredictSalience}(S_t)$\\
%        $E \gets \operatorname{APCluster}(P, S_t)$ \\
%        $U_t \gets \operatorname{SelectUpdates}(E,P,U)$\\
%        $U \gets U \cup U_t$
%}
% \caption{Temporal Summarization Algorithm}\label{alg:temporal-summarization}
%\end{algorithm}

@@

In order to fit a regression model, the sentences are represented as feature
vectors (described in the next section).
We fit a Gaussian process (GP) regression model to this data.
\cite{rasmussen:gaussian-process-book}.  GP regressors are a
class of data-driven, non-parametric model generalizing the multi-variate
Gaussian to the infinite dimensional setting.  
%GP's are general
%and are state of the art for many regression tasks.  
GP's are non-parametric and rely on a covariance matrix $\kernelMatrix$, measuring the similarity between pairs of instances, in our 
case sentences.  In our experiments, we used a radial basis 
function (RBF) kernel.  
Our features fall naturally into five groups and we use a separate RBF kernel
for each, using the sum of each feature group kernel as the final input
to the GP model.

@@

\subsubsection{Features}
We want our model to be predictive across different event instances so we avoid lexical features.  Instead, we extract a variety of features including language model scores, geographic relevance, and temporal relevance from each sentence.

\paragraph{Basic Features}
%KM - Would be good to have quick justification of these features. I added a
%sentence. Feel free to edit or remove.

We employ several basic features that have been used previously in supervised models to rank sentence salience \cite{kupiec1995trainable,conroy2001using}. These include sentence length, the number of capitalized words normalized by sentence length, document position, number of named entities.  
The data stream comprises text extracted from raw html documents;
these features help to downweight sentences that are not actually article 
content (e.g. web page titles, links to other content) or
more heavily weight important sentences (e.g., that appear in
prominent positions such as paragraph initial or article initial).

@@

\paragraph{Language Model Features}\label{subsubsec:lm}
Language models allow us to measure the likelihood of a sentence having been 
produced from a particular source.  We consider two types of language model 
features.  The first model is estimated from a corpus of generic news 
articles \ckcomment{(we used the 199?-200? Associated Press and New York Times sections of the Gigaword corpus)}.  
This model is intended to assess the general writing quality (grammaticality, word usage) of an input sentence and helps our model to select sentences
written in the newswire style.

@@

\fdcomment{stress here that this technique can be applied in situations beyond crisis; any situation where the
expected summary is similar to some previously seen target or some side information.}

These models are intended to detect sentences similar to those appearing in 
summaries of other events in the same category 
(e.g. most earthquake summaries are likely to include higher probability for 
ngrams including the token `magnitude').

@@

%For both models, we Finally, we extract the percentage of capitalized words,
%and sentence length as features. These last two features also help to
%identify sentences that are less likely to contain relevant content-- overly
%long and heavily capitalized sentences in our corpus were likely to be long
%strings of web-page headlines, section headers, and other irrelevant page
%structure.

\paragraph{Geographic Relevance Features}
\fdcomment{need introduction here.  what are geographic features?}
There are two challenges to using geographic features. First we do not 
know where the event is and second most sentences do not contain references
to a location.
We address the first issue by extracting all locations from sentences at the
current hour and looking up their latitude and 
longitude using a publicly available geolocation service. 
\fdcomment{is this all documents in the bucket or the filtered set?  is it a biased sample?}
Since the document
stream contains documents that are at least somewhat relevant to the event,
we assume in aggregate the locations should give us a rough area of interest.
The locations are clustered (using affinity propagation and uniform salience)
\fdcomment{affinity propagation undefined}
and we treat the resulting cluster centers
as the event locations for the current time.

To address the second issue, 
we compute geographic relevance features for the document as a whole and all
sentences from that document receive the same feature value.
Using the locations in each document, we compute the median distance to the 
nearest event location. Because document position is a good indicator 
of importance we also compute the distance of the first mentioned
location to the nearest event location. Because some events can move, we also
compute these distances to event locations from the previous hour.

\paragraph{Temporal Relevance Features}
\fdcomment{motivate burstiness; explain what it is before using the expression.}
Our data consists of hourly crawls of online content and so we exploit the temporality of corpus by capturing the burstiness of a sentence, i.e.  the change in word frequency from one hour to the next.``Bursty'' sentences often indicate new and important data.

We compute the IDF for each hour in our data stream. 
For each sentence, the average TFIDF for the current hour $t$ is taken as a 
feature. Additionally, we use the difference in average TFIDF from time $t$
to $t-i$ for $i = \{1, \ldots, 24\}$ to measure how the TFIDF scores for the 
sentence have changed over the last 24 hours. \fdcomment{unclear.  is it just changing the IDF part of the 
items in the average?}

The final temporal feature is the hours since the event started.
%difference 
%Using these IDF 
%calculations, the avg TFIDF is calculated for
%Let $D_t$ be the set of web pages at time $t$ and let $s = \{w_1,\ldots,w_n\}$ be a sentence from a page $d \in D_t$.  We calculate the 1-hour burstiness of sentence $s$ from document $d$ at hour $t$  as 
%\begin{align*}
%\operatorname{b}_1(s,d,t) = \frac{1}{|s|} \sum_{w \in s} \Bigg( &
%\operatorname{tf-idf}_t(w,d)  \\ & \left. - \frac{\sum_{d^\prime \in D_{t-1}:
%w \in d^\prime } \operatorname{tf-idf}_{t-1}(w,d^\prime)}{|\{d^\prime \in
%D_{t-1}: w \in d^\prime\}|} \right) \end{align*}

@@

We combine the output of our salience prediction model with the affinity
propagation algorithm to identify a set of exemplar sentences 
for each input batch. 
Affinity propagation (AP) is a clustering algorithm
that identifies a subset of datapoints as exemplars and forms clusters
by assigning the remaining points to one of the exemplars. AP attempts to 
maximize the net similarity objective 
\[ \mathcal{S} = \sum_{i : i \neq e_i}^n \operatorname{sim}(i,e_i) 
+ \sum_{i : i = e_i}^n \operatorname{pref}(e_i)  \]
where $e_i$ is the exemplar of the $i$-th data point, and functions
$\operatorname{sim}$ and $\operatorname{salience}$ express the pairwise 
similarity
of data points and the apriori preference of a data point to be an exemplar
respectively. 
AP differs from other $k$-centers algorithms in that it simultaneously 
considers all data points as exemplars, making it less prone to finding
local optima as a result of poor initialization. Furthermore, the 
second term in $\mathcal{S}$ incorporates the individual importance of 
data points as candidate exemplars; most other clustering algorithms only make
use of the first term, i.e. the pairwise similarities between data points.

>>>>

\input{nugget_examples.tex}

Throughout our treatment of our algorithm, the \emph{salience} 
of an update captures the degree to which it reflects an 
event's nuggets.  Assuming that we have a text representation 
of our nuggets, the salience of an update $u$  with respect to a set of nuggets $N$ is
defined as,
\begin{align}
\operatorname{salience}(u) = \operatorname{max}_{n \in N} 
\operatorname{sim}(u, n) \label{eq:salience}
\end{align}
where $\operatorname{sim}(\cdot)$ is the semantic similarity such as
the cosine similarity of latent vectors associated with the update and 
nugget text \cite{?}.

\subsection{Update Summarization}

Our system architecture follows a simple pipeline design where each
stage provides an additional level of processing or filtering of the input
sentences.
We begin with an empty update summary $U$.
At each hour we receive a new batch of sentences $b_t$ from the stream
and perform the following actions:
\begin{enumerate}
  \item predict the salience of sentences in $b_t$ (Section~\ref{sec:salpred}),
  \item select a set of exemplar sentences in $b_t$ by combining 
      clustering with 
      salience predictions (Section~\ref{sec:exsel}),
  \item add the most novel and salient exemplars 
      to $U$ (Section~\ref{sec:upsel}).
\end{enumerate}
The resultant list of updates $U$ is our summary of the event.

@@

In order to fit a regression model, the sentences are represented as feature
vectors (described in Section~\ref{sec:features}).
We fit a Gaussian process (GP) regression model to this data.
\cite{rasmussen:gaussian-process-book}.  GP regressors are a
class of data-driven, non-parametric model generalizing the multi-variate
Gaussian to the infinite dimensional setting.  
%GP's are general
%and are state of the art for many regression tasks.  
GP's are non-parametric and rely on a covariance matrix $\kernelMatrix$, measuring the similarity between pairs of instances, in our 
case sentences.  In our experiments, we used a radial basis 
function (RBF) kernel.  
Our features fall naturally into five groups and we use a separate RBF kernel
for each, using the sum of each feature group kernel as the final input
to the GP model.

@@

\subsubsection{Features}
\label{sec:features}
We want our model to be predictive across different event instances so we avoid lexical features.  Instead, we extract a variety of features including language model scores, geographic relevance, and temporal relevance from each sentence.

\paragraph{Basic Features}
We employ several basic features that have been used previously in supervised models to rank sentence salience \cite{kupiec1995trainable,conroy2001using}. These include sentence length, the number of capitalized words normalized by sentence length, document position, number of named entities.  
The data stream comprises text extracted from raw html documents;
these features help to downweight sentences that are not actually article 
content (e.g. web page titles, links to other content) or
more heavily weight important sentences (e.g., that appear in
prominent positions such as paragraph initial or article initial).

@@

\paragraph{Language Model Features}\label{subsubsec:lm}
Language models allow us to measure the likelihood of a sentence having been 
produced from a particular source.  We consider two types of language model 
features.  The first model is estimated from a corpus of generic news 
articles (we used the 1995-2010 Associated Press section of the 
Gigaword corpus).  
This model is intended to assess the general writing quality (grammaticality, word usage) of an input sentence and helps our model to select sentences
written in the newswire style.

@@

%\fdcomment{stress here that this technique can be applied in situations beyond crisis; any situation where the
%expected summary is similar to some previously seen target or some side information.}

These models are intended to detect sentences similar to those appearing in 
summaries of other events in the same category 
(e.g. most earthquake summaries are likely to include higher probability for 
ngrams including the token `magnitude'). While we focus our system on the 
language of news and disaster, we emphasize that the use of language modeling 
can be an effective feature for multi-document summarization for other 
domains that have related text corpora.

@@

\paragraph{Geographic Relevance Features}
%\fdcomment{need introduction here.  what are geographic features?}
The disasters in our corpus are all phenomena that
affect some part of the word. 
Where possible, we would like to capture a sentence's proximity to the event,
i.e. when a sentence references a location, it should be close to the 
area of the disaster.

There are two challenges to using geographic features. First we do not 
know where the event is and second most sentences do not contain references
to a location.
We address the first issue by extracting all locations from 
documents relevant to the event at the
current hour and looking up their latitude and 
longitude using a publicly available geolocation service. 
%\fdcomment{is this all documents in the bucket or the filtered set?  is it a biased sample?}
Since the documents that are at least somewhat relevant to the event,
we assume in aggregate the locations should give us a rough area of interest.
The locations are clustered %(using affinity propagation and uniform salience)
%\fdcomment{affinity propagation undefined}
and we treat the resulting cluster centers
as the event locations for the current time.

To address the second issue, 
we compute features for the document as a whole and all
sentences from that document receive the same feature value.

Using the locations in each document, we compute the median distance to the 
nearest event location. Because document position is a good indicator 
of importance we also compute the distance of the first mentioned
location to the nearest event location. All sentences in the document take
as features these two distance calculations.
Because some events can move, we also
compute these distances to event locations from the previous hour.

\paragraph{Temporal Relevance Features}
%\fdcomment{motivate burstiness; explain what it is before using the expression.}
As we track events over time, it is likely that the coverage of the event 
may die down, only to spike back up when there is a breaking development.
Alternatively, words that were salient initially may become background context,
and over time lose their importance.
Identifying terms that are ``bursty,'' i.e. suddenly peaking in usage,
can help to locate novel sentences that are part of the most recent reportage
and have yet to fall into the background.

We compute the IDF for each hour in our data stream. 
For each sentence, the average TFIDF for the current hour $t$ is taken as a 
feature. Additionally, we use the difference in average TFIDF from time $t$
to $t-i$ for $i = \{1, \ldots, 24\}$ to measure how the TFIDF scores for the 
sentence have changed over the last 24 hours, i.e. we keep the sentence
term frequencies fixed and compute the difference in IDF. Large changes
in the is value indicate the sentence contains bursty terms.
%\fdcomment{unclear.  is it just changing the IDF part of the 
%items in the average?}
We also use the time (in hours) since the event started as a feature.
%difference 
%Using these IDF 
%calculations, the avg TFIDF is calculated for
%Let $D_t$ be the set of web pages at time $t$ and let $s = \{w_1,\ldots,w_n\}$ be a sentence from a page $d \in D_t$.  We calculate the 1-hour burstiness of sentence $s$ from document $d$ at hour $t$  as 
%\begin{align*}
%\operatorname{b}_1(s,d,t) = \frac{1}{|s|} \sum_{w \in s} \Bigg( &
%\operatorname{tf-idf}_t(w,d)  \\ & \left. - \frac{\sum_{d^\prime \in D_{t-1}:
%w \in d^\prime } \operatorname{tf-idf}_{t-1}(w,d^\prime)}{|\{d^\prime \in
%D_{t-1}: w \in d^\prime\}|} \right) \end{align*}

@@

We combine the output of our salience prediction model with the affinity
propagation algorithm to identify a set of exemplar sentences 
for each input batch. 
Affinity propagation (AP) is a clustering algorithm
that identifies a subset of datapoints as exemplars and forms clusters
by assigning the remaining points to one of the exemplars. AP attempts to 
maximize the net similarity objective 
\[ \mathcal{S} = \sum_{i : i \neq e_i}^n \operatorname{sim}(i,e_i) 
+ \sum_{i : i = e_i}^n \operatorname{salience}(e_i)  \]
where $e_i$ is the exemplar of the $i$-th data point, and functions
$\operatorname{sim}$ and $\operatorname{salience}$ express the pairwise 
similarity
of data points and the apriori preference of a data point to be an exemplar
respectively. 
AP differs from other $k$-centers algorithms in that it simultaneously 
considers all data points as exemplars, making it less prone to finding
local optima as a result of poor initialization. Furthermore, the 
second term in $\mathcal{S}$ incorporates the individual importance of 
data points as candidate exemplars; most other clustering algorithms only make
use of the first term, i.e. the pairwise similarities between data points.

### 2015-02-27 results.tex
Latest revision ck
Latest rev ck

\subsection{ROUGE}

Table~\ref{tab:rouge} shows our results for the retrospective evaluation using ROUGE. 
Both versions of AP+Salience show improvement on ngram precision and recall over the 
baselines. This improvement is statistically significant at the $\alpha = .01$ level
using the Wilcoxon signed-rank test. Additionally, AP+Salience's improvement over
AP+Salience TR in recall and F-measure is also statistically significant at the same level.

AP+Salience maintains its performance above the baselines over time as well. Table~\ref{fig:trouge}
shows the ROUGE-1 performance over time. Within the initial days of the event, recall is 
competitive amongst system. However, both AP+Salience variants are able to achieve higher
precision early on.

\subsection{Nugget Evaluation}

\subsection{Feature Ablation}

\begin{table}[h]
\centering
% centering table
\begin{tabular}{l c c c}
% creating 10 columns
\multicolumn{4}{c}{ROUGE-1}\\
\hline
\hline
% inserting double-line
$\mathrm{System}$ & $\mathrm{Prec.}$ & $\mathrm{Recall}$ & $\mathrm{F}_1$\\
[0.5ex]
\hline
AP+Salience &          0.289   &    0.344  &  0.312  \\
AP+Salience TR &       0.278   &    0.342  &  0.303 \\
AP    &         0.246   &    0.285  &  0.263 \\
HAC     &        0.238   &    0.276  &  0.255 \\
Rank    &        0.229   &    0.271  &  0.247 \\
% inserts single-line
% Entering 1st row
%& &soft &1 & $-1$ & 1 & 1 & $-1$ & $-1$ & 1\\
%[-1ex]
%\raisebox{1.5ex}{Police} &\raisebox{1.5ex}{5}&hard& 2 & $-4$ & 4 & 4 & $-2$ & $-4$ & 4\\
%[1ex]
% Entering 2nd row
%& &soft & 1 & $-1$ & 1 & 1 & $-1$ & $-1$ & 1\\
%[-1ex]
%\raisebox{1.5ex}{Beethoven}&\raisebox{1.5ex}{5}& hard&8 & $-8$ & 2 & 8 & $-8$ & $-8$ & 6
%\\
%[1ex]
% Entering 3rd row
%& &soft & 1 & $-1$ & 1 & 1 & $-1$ & $-1$ & 1
%\\
%[-1ex]
%\raisebox{1.5ex}{Metallica} &\raisebox{1.5ex}{5}& hard&4 & $-8$ & 8 & 4 & $-8$ & $-8$ & 8\\
%[1ex]
% [1ex]adds vertical space
\hline % inserts single-line
\end{tabular}
~\\[1ex]
~\\
\begin{tabular}{l c c c}
% creating 10 columns
\multicolumn{4}{c}{ROUGE-2}\\
\hline
\hline
% inserting double-line
$\mathrm{System}$ & $\mathrm{Prec.}$ & $\mathrm{Recall}$ & $\mathrm{F}_1$\\[0.5ex]
\hline
AP+Salience & 0.047 &     0.056 &   0.051 \\
AP+Salience TR & 0.044  &    0.054 &  0.047 \\
AP & 0.033  &    0.038 &  0.035 \\
HAC & 0.031  &    0.035 &  0.034 \\
Rank & 0.031  &    0.037 &  0.034 \\
\hline % inserts single-line
\end{tabular}
\caption{System ROUGE performance.} % title name of the table
\label{tab:rouge}
\end{table}

\begin{figure*}
  \includegraphics[width=\textwidth]{rouge-time.png}
\caption{System ROUGE-1 performance over time.}
\label{fig:trouge}
\end{figure*}

>>>>

\subsection{ROUGE}
\input{rouge_cum.tex}

Table~\ref{tab:rouge} shows our results for the retrospective evaluation using ROUGE. 
AP+Salience shows improvement on ngram precision and recall over the 
baselines. This improvement is statistically significant for all ngram and longest common substring
precision, recall, and F-measure
at the $\alpha = .01$ level
using the Wilcoxon signed-rank test.

AP+Salience maintains its performance above the baselines over time as well. Figure~\ref{fig:trouge}
shows the ROUGE-1 scores over time. 
We emphasize the difference in unigram precision (bigram precision is not shown but it follows
similar curve).
Within the initial days of the event, AP+Salience is able to take the lead over the over 
systems in ngram precision. 
The AP+Salience model is better able to find salient updates
earlier on; for the disaster domain, this is an especially important quality of the model.

Moreover, the AP+Salience's recall is not diminished by the high precision and remains competitive with AP.
Over time AP+Salience's recall also begins to pull away, while the other models start to suffer
from topic drift.

\subsection{Expected Gain and Comprehensiveness}

Figure~\ref{fig:nperf} shows the expected gain across a range of 
similarity thresholds, where thresholds closer to 1 are more conservative
estimates. The ranking of the systems remains constant across the 
sweep with 
AP+Salience beating all baseline systems.
Predicting salience in general is helpful for keeping a summary on topic as
the RS approach out performs the clustering only approaches
on expected gain.

When looking at the comprehensiveness of the summaries AP outperforms
 AP+Salience. The compromise encoded in the AP+Salience objective
function, between being representitive and being salient, is seen clearly here
where the performance of the AP+Salience methods is lower bounded by 
the salience focused RS
system and upper bounded by the clustering only AP system.
Overall, AP+Salience achieves the best balance of these two metrics.

\subsection{Feature Ablation}
\input{rouge_fa.tex}

Table~\ref{tab:farouge} shows the results of our feature ablation
tests. Removing the language models yields a statistically 
significant drop in both ngram recall and F-measure. 
Interestingly, removing the basic features leads to an
increase in both unigram and bigram precision; in the bigram
case this is enough to cause a statistically significant increase
in F-measure over the full model. In other words, the generic features
actually lead to an inferior model when we can incorporate more appropriate
domain specific features. \ckcomment{Kathy: maybe we should mention Spark Jones here about the idea that there is no such thing as generic summarization?}

\begin{figure}
    \includegraphics[]{rouge-time.eps}
\caption{System ROUGE-1 performance over time.}
\label{fig:trouge}
\end{figure}

\begin{figure}
  \includegraphics[]{nuggets-metrics.eps}
\caption{Avg. Gain and Comprehensiveness performance.}
\label{fig:nperf}
\end{figure}
